---
title: "Bringing it All Together"
description: |
  Some examples of using all my tools from EDS 221 to wrangle and create data visualizations.
author:
  - name: Jake Eisaguirre
    url: {}
date: 08-20-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, echo = FALSE, include=F}

knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
library(lubridate)
library(kableExtra)
```


```{r}

fids <- read_csv(here("_posts","2021-08-20-bringing-it-all-together", "HTL-MAR-FiddlerCrabBodySize_copy.csv"))

```


## Data Exploration



```{r}
ggplot(fids, aes(x = Latitude, y = carapace_width)) + 
  geom_point(aes(color = Latitude)) +
  labs(title = "Carapace Width by Latitude",
       x = "Latitude", 
       y = "Carapace Width (mm)", 
       caption = "Figure A is the observed Caprapace widths at latitude degrees") +
  theme_classic()
  
```


Here we have a nice figure in which we can do some initial data exploration regarding carapace width by latitude.


## Data Table



```{r}

data <- fids %>%
  clean_names() %>% 
  group_by(site) %>% 
  summarize(mean_carapace_width = round(mean(carapace_width), 2), sd_carapace_width = round(sd(carapace_width), 2), sample_size = n())

kable(data, col.name = c("Site", "Mean Carapace Width (mm)", "SD Carapace Width (mm)", "Sample Size")) %>% 
  kable_classic()
  

```


Here is a data table with some basic info about our fiddler crabs at different sites


## Temperature vs Carapace Width



```{r, fig.alt = "Scatterplot of mean annual temperature compared to mean carapace width. Increases in temeperature result in decreases in mean width size."}
mean_data = fids %>% 
  clean_names() %>%
  group_by(matw) %>% 
  summarize(mean_carapace_width = mean(carapace_width))
  
ggplot(mean_data, aes(x = matw, y = mean_carapace_width)) + 
  geom_point() + 
  labs(title = "Mean Carapace Width by Temp", 
       x = "Mean Annual Water Temperature", 
       y = "Mean Carapace Width (mm)", 
       caption = "Figure B shows a negative trend in mean carapace width with increase in water temperature.") +
  theme_classic() +
  geom_smooth(method = "lm")

```


Here we can see the possible correlation between mean carapace width and temperature



### Citations 
* Johnson, D. 2019. Fiddler crab body size in salt marshes from Florida to Massachusetts, USA at PIE and VCR LTER and NOAA NERR sites during summer 2016. ver 1. Environmental Data Initiative. https://doi.org/10.6073/pasta/4c27d2e778d3325d3830a5142e3839bb (Accessed 2021-08-20).

* Wickham et al., (2019). Welcome to the tidyverse. Journal of Open Source Software, 4(43), 1686,
  https://doi.org/10.21105/joss.01686

* Kirill Müller (2020). here: A Simpler Way to Find Your Files. R package version 1.0.1.
  https://CRAN.R-project.org/package=here
  
* Garrett Grolemund, Hadley Wickham (2011). Dates and Times Made Easy with lubridate. Journal of Statistical Software, 40(3),
  1-25. URL https://www.jstatsoft.org/v40/i03/.

* Sam Firke (2021). janitor: Simple Tools for Examining and Cleaning Dirty Data. R package version 2.1.0.
  https://CRAN.R-project.org/package=janitor

* Hao Zhu (2021). kableExtra: Construct Complex Table with 'kable' and Pipe Syntax. R package version 1.3.4.
  https://CRAN.R-project.org/package=kableExtra
```