---
title: "Using Data Science to Catch Fish"
description: |
  The free webstie I was using to find temperature breaks and chlorophyll levels in our local waters was not accurate. So I created my own figures to help determine where to find game fish.  
author:
  - name: Jake Eisaguirre
    url: {}
date: 11-08-2021
output:
  distill::distill_article:
    self_contained: false
    code_folding: hide
---

### Shiny App link: http://chuna.shinyapps.io/Chuna_Findaa (still debugging deployment)


### Update: I have now completed the list of items mentioned below. I am now attempting to publish this work on an active site in order to help my fellow local so-cal anglers catch more fish. I am now working to incorporate predicted swell models from CDIP buoys. See bottom of post for most current map.   

```{r setup, include=FALSE, collapse=TRUE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(here)
library(tidyverse)
library(ggExtra)
library(patchwork)
library(gridExtra)
library(rerddapXtracto)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(raster)
library(leaflet)
library(fasterize)
library(rnoaa)
library(htmlwidgets)
library(RColorBrewer)
library(mapview)
library(leafem)
library(colorRamps)




```

### Below is my first attempt to visualize SST and Chlorophyll.

##### Initial Goals: This is a work in progress. My next challenge is to include bathymetry lines. Next I would like to add bouy locations that gives live readings of sea conditions. Then I want to use the `leaflet` package to overlay the two figures so I can toggle between SST and Chlorophyll. Finally I would like to make the figure interacive so one can zoom in and out. 

```{r}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(tidyverse)
library(ggExtra)
library(patchwork)
library(gridExtra)
library(rerddapXtracto)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(raster)
library(leaflet)
library(colorspace)




```

#### Read in SST and Chloro from NOAA Aqua MODIS Satelite
```{r}
require("rerddap")

raw_sst <- griddap('erdMWsstd1day_LonPM180',
 time = c('last','last'),
 latitude = c(33.1125, 34.9),
 longitude = c(-120.69999999999999, -118.75),
 fmt = "csv")




raw_chloro <- griddap('erdMWchla1day_LonPM180',
 time = c('last','last'),
 latitude = c(33.1125, 34.9),
 longitude = c(-120.69999999999999, -118.75),
 fmt = "csv")


```

#### Read in CA and N. Channel Islands outline
```{r}


cha <- read_sf(here("_posts", "2021-09-23-sst", "shape", "channel_islands.shp")) %>% 
  mutate(geometry = st_transform(geometry, "+proj=longlat +ellps=WGS84 +datum=WGS84")) %>% 
  filter(!NAME %in% c("Santa Catalina", "San Clemente"))

ca <- ne_countries(country = "united states of america", scale = "large", returnclass = "sf")

```

#### Clean SST data
```{r}

sst <- raw_sst[-c(2)] 
  
clean_sst <- sst[-c(1),] 

mut_sst <- clean_sst %>% 
  mutate(data.frame(longitude = as.numeric(clean_sst$longitude))) %>% 
  mutate(data.frame(latitude = as.numeric(clean_sst$latitude))) %>% 
  mutate(data.frame(sst = as.numeric(clean_sst$sst))) %>% 
  filter(sst > 0)

final_sst <- mut_sst %>% 
  mutate(sst = (sst * (9/5) + 32 )) %>% 
  mutate(sst = (sst - 3))

```

#### SST plot
```{r}
sst_plot <- ggplot() +
  geom_raster(data = final_sst, aes(y=latitude, x=longitude, fill = sst), 
              show.legend = T, 
              interpolate = T, 
              hjust = 1, 
              vjust = 1) +
  scale_fill_viridis_c(option = "plasma", begin = 0) +
  theme_light() +
  labs(x = "Longitude",
       y = "Latitude",
       #title = "3 Day Composite",
       fill = "Sea Surface 
Temperature (F)") +
  theme(panel.grid = element_line(F),
        legend.position = "bottom",
        panel.background = element_rect(fill = "grey70"),
        axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),
        plot.margin = unit(c(0,0,0,0), "inches")) +
  scale_x_continuous(breaks = c(-120.5, -120.25, -120, -119.75, -119.5, -119.25, -119), 
                     expand = c(0,0)) +
  scale_y_continuous(breaks = c(33.25, 33.5, 33.75, 34, 34.25, 34.5),
                                        expand = c(0,0)) +
  geom_sf(data = cha, color = "black", fill = "grey60") +
  geom_sf(data = ca, color = "black", fill = "grey60") +
  coord_sf(xlim = c(-120.69999999999999, -118.75), ylim = c(33.1125, 34.9), expand = FALSE) 

```

#### Clean Chloro data
```{r}

chloro <- raw_chloro[-c(2)] 
  
clean_chloro <- chloro[-c(1),] 

final_chloro <- clean_chloro %>% 
  mutate(data.frame(longitude = as.numeric(clean_chloro$longitude))) %>% 
  mutate(data.frame(latitude = as.numeric(clean_chloro$latitude))) %>% 
  mutate(data.frame(chlorophyll = as.numeric(clean_chloro$chlorophyll))) %>% 
  filter(chlorophyll < 3 )


```

#### Chloro plot
```{r}

chloro_plot <- ggplot() +
  geom_raster(data = final_chloro, aes(y=latitude, x=longitude, fill = chlorophyll), 
              show.legend = T, 
              interpolate = T, 
              hjust = 1, 
              vjust = 1) +
  scale_fill_viridis_c(option = "turbo") +
  theme_light() +
  labs(x = "Longitude",
       y = "",
       fill = "Chlorophyll
(mg m^-3)") +
  theme(panel.grid = element_line(F),
        panel.background = element_rect(fill = "grey70"),
        legend.position = "bottom",
        axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        plot.margin = unit(c(0,0.8,0,0), "inches")) +
  scale_x_continuous(breaks = c(-120.5, -120.25, -120, -119.75, -119.5, -119.25, -119), 
                     expand = c(0,0)) +
  scale_y_continuous(breaks = c(33.25, 33.5, 33.75, 34, 34.25, 34.5), 
                                expand = c(0,0), 
                     position = "right") +
  geom_sf(data = cha, color = "black", fill = "grey60") +
  geom_sf(data = ca, color = "black", fill = "grey60") +
  coord_sf(xlim = c(-120.69999999999999, -118.75), ylim = c(33.1125, 34.9), expand = FALSE)
  
  
```  
#### Final Product
```{r}
grid.arrange(sst_plot, chloro_plot, ncol=2)
```

#### Read in SST and Chloro from Aqua Modis Sat
```{r}

date <- as.Date(Sys.time()) -1
  
date <- as.Date(date) %>% 
  paste0("T12:00:00Z")

past_date <- as.Date(date) -1 

past_date <- as.Date(past_date) %>% 
  paste0("T12:00:00Z") 

url <- paste0("https://thredds.jpl.nasa.gov/thredds/ncss/OceanTemperature/MUR-JPL-L4-GLOB-v4.1.nc?var=analysed_sst&north=35.0&west=-121.5&east=-116.5&south=31.75&disableLLSubset=on&disableProjSubset=on&horizStride=1&time_start=", past_date, "&time_end=", date, "&timeStride=1&addLatLon=true")

sst_data <- download.file(url, here("sst.nc"))

sst_ras <- here("sst.nc")


chl_date <- as.Date(Sys.time()) -2

chl_date <- as.Date(chl_date) %>% 
  paste0("T12:00:00Z") 

url_1 <- paste0("https://coastwatch.pfeg.noaa.gov/erddap/griddap/erdMWchla3day_LonPM180.nc?chlorophyll%5B(", chl_date,")%5D%5B(0.0)%5D%5B(31.75):(35.0)%5D%5B(-121.5):(-116.5)%5D&.draw=surface&.vars=longitude%7Clatitude%7Cchlorophyll&.colorBar=%7C%7C%7C%7C%7C&.bgColor=0xffccccff")

download.file(url_1, here("chl_1.nc"))

chl_ras <- here("chl_1.nc")



so_cal_bath <- marmap::getNOAA.bathy(
  lon1 = -116.5,
  lon2 = -121.5,
  lat1 = 35.0,
  lat2 = 31.75,
  resolution = 1) %>% 
  marmap::as.raster() %>% 
  raster::rasterToPoints() %>% 
  base::as.data.frame() %>% 
  filter(layer <= 0) %>% 
  mutate(layer = layer*-1)

```

#### Read in Shape Files
```{r}


cha <- read_sf(here("_posts", "2021-09-23-sst", "shape", "channel_islands.shp")) %>% 
  mutate(geometry = st_transform(geometry, "+proj=longlat +ellps=WGS84 +datum=WGS84")) %>% 
  filter(!NAME %in% c("Santa Catalina", "San Clemente"))

ca <- read_sf(here("_posts", "2021-09-23-sst", 
                   "s_11au16", "s_11au16.shp")) %>% 
  mutate(geometry = st_transform(geometry, "+proj=longlat +ellps=WGS84 +datum=WGS84")) %>% 
  filter(NAME == "California")

mpa <- read_sf(here("_posts", "2021-09-23-sst", "ds582", "ds582.shp")) %>%   
  mutate(geometry = st_transform(geometry, "+proj=longlat +ellps=WGS84 +datum=WGS84")) %>% 
  filter(Study_Regi == "SCSR")

smr <- read_sf(here("_posts", "2021-09-23-sst", "ds582", "ds582.shp")) %>%   
  mutate(geometry = st_transform(geometry, "+proj=longlat +ellps=WGS84 +datum=WGS84")) %>% 
  filter(Study_Regi == "SCSR") %>% 
  filter(Type %in% c("SMR", "FMR"))

smca <- read_sf(here("_posts", "2021-09-23-sst", "ds582", "ds582.shp")) %>%   
  mutate(geometry = st_transform(geometry, "+proj=longlat +ellps=WGS84 +datum=WGS84")) %>% 
  filter(Study_Regi == "SCSR") %>% 
  filter(Type == "SMCA")

no_take <- read_sf(here("_posts", "2021-09-23-sst", "ds582", "ds582.shp")) %>%   
  mutate(geometry = st_transform(geometry, "+proj=longlat +ellps=WGS84 +datum=WGS84")) %>% 
  filter(Study_Regi == "SCSR") %>% 
  filter(Type == "SMCA (No-Take)")

merged_shapes_mask <- bind_rows(ca, cha)


```


#### Create Rasters and Masks
```{r}
#sst raster
ras_sst <- raster(sst_ras)

ras_sst <- ((9/5) * (ras_sst$analysed.sea.surface.temperature - 273) + 32)

new_ras_sst <- raster(xmn = -121.5,
                      xmx = -116.5,
                      ymn = 31.0,
                      ymx = 35.0,
                      res = c(0.002, 0.002))


re_samp_sst <- resample(ras_sst, new_ras_sst, method = "bilinear")


cropped_sst <- mask(re_samp_sst, merged_shapes_mask, inverse = T)


final_re_samp_sst <- projectRaster(cropped_sst, crs = 4326) #final re-sammpled SST raster


# Chloro Raster
ras_chl <- raster(chl_ras)

ras_chl <- abs(log(ras_chl$Concentration.Of.Chlorophyll.In.Sea.Water))

new_ras_chl <- raster(xmn = -121.5,
                      xmx = -116.5,
                      ymn = 31.0,
                      ymx = 35.0,
                      res = c(0.002, 0.002))

re_samp_chl <- resample(ras_chl, new_ras_chl, method = "bilinear")

cropped_chl <- mask(re_samp_chl, merged_shapes_mask, inverse = T)

re_samp_chl <- projectRaster(cropped_chl, crs = 4326) #final re-sampled Chloro raster


# bath Raster

bath_ras <- rasterFromXYZ(so_cal_bath, crs = 4326)

new_ras_bath <- raster(xmn = -121.5,
                      xmx = -116.5,
                      ymn = 31.0,
                      ymx = 35.0,
                      res = c(0.002, 0.002))


re_samp_bath <- resample(bath_ras, new_ras_bath, method = "bilinear") #final re-sampled Bathy raster

cropped_bath <- mask(re_samp_bath, merged_shapes_mask, inverse = T)

re_samp_bath <- projectRaster(cropped_bath, crs = 4326) #final re-sampled Chloro raster
```

#### Color pals
```{r}
r_chl <- as.data.frame(rasterToPoints(ras_chl)) %>% 
  rename(chlorophyll = layer)
chl_pal <- colorNumeric(palette =colorspace::sequential_hcl(25,
  h = c(300, 75), c = c(35, 95), l = c(15, 90), power = c(0.8, 1.2)), domain = r_chl$chlorophyll)


rev <- rev(sequential_hcl(40,
  h = c(260, 220), c = c(74, 112, 39), l = c(17, 88), power = c(0, 1.3)))
bath_pal <- colorNumeric(palette = rev, domain = so_cal_bath$layer)


r_sst <- as.data.frame(rasterToPoints(ras_sst)) %>% 
  rename(sst = analysed.sea.surface.temperature)
sst_pal <- colorNumeric(palette = matlab.like(25), domain = r_sst$sst)

#Buoy Icon
Buoy <-makeIcon('icons8-buoy-50.png', iconWidth = 30, iconHeight = 30)

```

#### Bluild leaflet
```{r}
sst_leaf <- leaflet(options = leafletOptions(minZoom = 8.5)) %>%
  addPolygons(data = cha, color = 'black', opacity = 1, weight = 2, fill = F) %>% 
  addPolygons(data = ca, color = 'black', opacity = 1, weight = 2, fill = F) %>% 
  addPolygons(data = smr, group = "MPA Status", color = "red", popup = ~NAME) %>%
  addPolygons(data = smca, group = "MPA Status", color = "blue", popup = ~NAME) %>% 
  addPolygons(data = no_take, group = "MPA Status", color = "purple", popup = ~NAME) %>% 
  addProviderTiles("Esri.OceanBasemap") %>%
  addRasterImage(x = final_re_samp_sst, colors = sst_pal, opacity = 0.7,
                 group = "Sea Surface Temp") %>%
  addRasterImage(x = re_samp_chl, colors = chl_pal, opacity = 0.7,
                 group = "Chlorophyll") %>%
  addRasterImage(x = re_samp_bath, colors = bath_pal, opacity = 0.7,
                 group = "Bathymetry") %>% 
  addMarkers(lng=-120.47, lat=34.273, 
            popup="<a href =https://www.ndbc.noaa.gov/station_page.php?station=46054>
                   West Buoy Data</a>", 
            group = "Buoys", 
            icon = Buoy) %>% 
  addMarkers(lng=-119.839, lat=34.241, 
            popup="<a href = https://www.ndbc.noaa.gov/station_page.php?station=46053>
                   East Buoy Data</a>", 
            group = "Buoys", 
            icon = Buoy) %>% 
  addMarkers(lng=-119.044, lat=33.758, 
            popup="<a href = https://www.ndbc.noaa.gov/station_page.php?station=46025>
                   Santa Monica Basin Buoy Data</a>", 
            group = "Buoys", 
            icon = Buoy) %>%
  addMarkers(lng=-119.565, lat=33.769, 
            popup="<a href = https://www.ndbc.noaa.gov/station_page.php?station=46251>
                   Santa Cruz Basin Buoy Data</a>", 
            group = "Buoys", 
            icon = Buoy) %>%
  addMarkers(lng=-120.213, lat=33.677, 
            popup="<a href =https://www.ndbc.noaa.gov/station_page.php?station=46069>
                   South Santa Rosa Buoy Data</a>", 
            group = "Buoys", 
            icon = Buoy) %>%
  addMarkers(lng=-118.641, lat=33.860, 
            popup="<a href = https://www.ndbc.noaa.gov/station_page.php?station=46221>
                   Santa Monica Bay Buoy Data</a>", 
            group = "Buoys", 
            icon = Buoy) %>%
  addMarkers(lng=-118.317, lat=33.618, 
           popup="<a href = https://www.ndbc.noaa.gov/station_page.php?station=46222>
                   San Pedro Buoy Data</a>", 
           group = "Buoys", 
           icon = Buoy) %>%
  addMarkers(lng=-118.181, lat=33.576, 
            popup="<a href = https://www.ndbc.noaa.gov/station_page.php?station=46253>
                   San Pedro South Buoy Data</a>", 
            group = "Buoys", 
            icon = Buoy) %>% 
  addMarkers(lng=-117.472, lat=33.178, 
              popup="<a href=https://www.ndbc.noaa.gov/station_page.php?station=46224>
                   Oceanside Offshore Buoy Data</a>", 
             group = "Buoys", 
             icon = Buoy) %>%
  addMarkers(lng=-117.391, lat=32.933, 
            popup="<a href = https://www.ndbc.noaa.gov/station_page.php?station=46225>
                   Torrey Pines Outer Buoy Data</a>", 
            group = "Buoys", 
            icon = Buoy) %>%
  addMarkers(lng=-117.501, lat=32.752, 
              popup="<a href=https://www.ndbc.noaa.gov/station_page.php?station=46258>
                   Mission Bay West Buoy Data</a>", 
             group = "Buoys", 
             icon = Buoy) %>%
  addMarkers(lng=-117.425, lat=32.517, 
             popup="<a href =https://www.ndbc.noaa.gov/station_page.php?station=46232>
                   Point Loma South Buoy Data</a>", 
             group = "Buoys", 
             icon = Buoy) %>%
  addMarkers(lng=-118.052, lat=32.499, 
            popup="<a href = https://www.ndbc.noaa.gov/station_page.php?station=46086>
                   San Clemente Basin Buoy Data</a>", 
            group = "Buoys", 
            icon = Buoy) %>%
  addMarkers(lng=-119.525, lat=32.388, 
            popup="<a href = https://www.ndbc.noaa.gov/station_page.php?station=46047>
                   Tanner Bank Buoy Data</a>", 
            group = "Buoys", 
            icon = Buoy,
            markerOptions(interactive = T, clickable = T, riseOnHover = T)) %>%
  addLegend(data = r_sst, pal = sst_pal, title = 'Sea Surface Temp', 
            position = "bottomright", values = ~sst, 
            opacity = 1, group = "Sea Surface Temp",
            bins = 8) %>% 
  addLegend(data = so_cal_bath, pal = bath_pal, title = 'Bathymetry', 
            position = "bottomright", 
            values = ~layer, opacity = 1, group = "Bathymetry") %>%
  addLegend(data = r_chl, pal = chl_pal, title = 'Chlorophyll', 
            position = "bottomright", 
            values = ~chlorophyll, opacity = 1, group = "Chlorophyll") %>% 
  addLayersControl(
    baseGroups = c("Sea Surface Temp", "Chlorophyll", "Bathymetry"),
    overlayGroups = c("Buoys", "MPA Status"),
    options = layersControlOptions(collapsed = FALSE)) %>% 
   htmlwidgets::onRender("
    function(el, x) {
      var updateLegend = function () {
          var selectedGroup = document.querySelectorAll('input:checked')[0].nextSibling.innerText.substr(1);

          document.querySelectorAll('.legend').forEach(a => a.hidden=true);
          document.querySelectorAll('.legend').forEach(l => {
            if (l.children[0].children[0].innerText == selectedGroup) l.hidden=false;
          });
      };
      updateLegend();
      this.on('baselayerchange', e => updateLegend());
    }") %>% 
  setView(lng = -119.200336, lat = 33.808464, zoom = 8.5) %>% 
  setMaxBounds(lng1 = -121.6,
               lat1 = 35.0,
               lng2 = -116.5,
               lat2 = 31.75) %>% 
  addMeasure(
    position = "bottomleft",
    primaryLengthUnit = "feet",
    primaryAreaUnit = "sqfeet",
    activeColor = "#3D535D",
    completedColor = "#7D4479") %>% 
 addMouseCoordinates()

sst_leaf
  

#saveWidget(app, file= "app.html", selfcontained = T)


```
#### Shiny app
```{r}

library(shiny)
ui <- fluidPage(
  
  titlePanel("So-Cal Fish Bite"),
  
  leafletOutput("sst_leaf", height = 700, width = "100%"),
  
  
  
)

server <- function(input, output, session){
  
  #Reactive
    sst_reactive <- reactive({
       final_re_samp_sst
    })
    



  #Map
  output$sst_leaf <- renderLeaflet({
    
    sst_leaf 
      

    
    })
  
}

shinyApp(ui = ui, server = server)

```






