---
title: "Using Data Science to Catch Fish"
description: |
  The free webstie I was using to find temperature breaks and chlorophyll levels in our local waters was not accurate. So I created my own figures to help determine where to find game fish.  
author:
  - name: Jake Eisaguirre
    url: {}
date: 09-23-2021
output:
  distill::distill_article:
    self_contained: false
---

##### This is a work in progress. My next challenge is to include bathymetry lines. Next I would like to add bouy locations that gives live readings of sea conditions. Then I want to use the `leaflet` package to overlay the two figures so I can toggle between SST and Chlorophyll. Finally I would like to make the figure interacive so one can zoom in and out. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(tidyverse)
library(ggExtra)
library(patchwork)
library(gridExtra)
library(rerddapXtracto)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(raster)
library(leaflet)




```

# Read in SST and Chloro from NOAA Aqua MODIS Satelite
```{r}
require("rerddap")

raw_sst <- griddap('erdMWsstd1day_LonPM180',
 time = c('last','last'),
 latitude = c(33.1125, 34.9),
 longitude = c(-120.69999999999999, -118.75),
 fmt = "csv")



three_day_chloro <- info('erdMWchla1day_LonPM180')
three_day_chloro

raw_chloro <- griddap('erdMWchla1day_LonPM180',
 time = c('last','last'),
 latitude = c(33.1125, 34.9),
 longitude = c(-120.69999999999999, -118.75),
 fmt = "csv")


# Buoy df

b_lat <-c(34.273, 34.241, 33.758, 33.769, 33.677)

b_long <- c(120.47, 119.839, 119.044, 119.565, 120.213)

b_name <- c("West Buoy", "East Buoy", "Santa Monic Baisin Buoy", "SCRIPPS", "BACK")

b_df <- data.frame(b_name, b_lat, b_long)


```

# Read in CA and N. Channel Islands outline
```{r}


cha <- read_sf(here("/Users","JaketheBoss", "Documents", "Bren", "SST Code", "SST", "SST", "data", "shape", "channel_islands.shp")) %>% 
  mutate(geometry = st_transform(geometry, "+proj=longlat +ellps=WGS84 +datum=WGS84")) %>% 
  filter(!NAME %in% c("Santa Catalina", "San Clemente"))

ca <- ne_countries(country = "united states of america", scale = "large", returnclass = "sf")

```

# Clean SST data
```{r}

sst <- raw_sst[-c(2)] 
  
clean_sst <- sst[-c(1),] 

mut_sst <- clean_sst %>% 
  mutate(data.frame(longitude = as.numeric(clean_sst$longitude))) %>% 
  mutate(data.frame(latitude = as.numeric(clean_sst$latitude))) %>% 
  mutate(data.frame(sst = as.numeric(clean_sst$sst))) %>% 
  filter(sst > 0)

final_sst <- mut_sst %>% 
  mutate(sst = (sst * (9/5) + 32 )) %>% 
  mutate(sst = (sst - 3))

```

# SST plot
```{r}
sst_plot <- ggplot() +
  geom_raster(data = final_sst, aes(y=latitude, x=longitude, fill = sst), 
              show.legend = T, 
              interpolate = T, 
              hjust = 1, 
              vjust = 1) +
  scale_fill_viridis_c(option = "plasma", begin = 0) +
  theme_light() +
  labs(x = "Longitude",
       y = "Latitude",
       #title = "3 Day Composite",
       fill = "Sea Surface 
Temperature (F)") +
  theme(panel.grid = element_line(F),
        legend.position = "bottom",
        panel.background = element_rect(fill = "grey70"),
        axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),
        plot.margin = unit(c(0,0,0,0), "inches")) +
  scale_x_continuous(breaks = c(-120.5, -120.25, -120, -119.75, -119.5, -119.25, -119), 
                     expand = c(0,0)) +
  scale_y_continuous(breaks = c(33.25, 33.5, 33.75, 34, 34.25, 34.5),
                                        expand = c(0,0)) +
  geom_sf(data = cha, color = "black", fill = "grey60") +
  geom_sf(data = ca, color = "black", fill = "grey60") +
  coord_sf(xlim = c(-120.69999999999999, -118.75), ylim = c(33.1125, 34.9), expand = FALSE) 

```

# Clean Chloro data
```{r}

chloro <- raw_chloro[-c(2)] 
  
clean_chloro <- chloro[-c(1),] 

final_chloro <- clean_chloro %>% 
  mutate(data.frame(longitude = as.numeric(clean_chloro$longitude))) %>% 
  mutate(data.frame(latitude = as.numeric(clean_chloro$latitude))) %>% 
  mutate(data.frame(chlorophyll = as.numeric(clean_chloro$chlorophyll))) %>% 
  filter(chlorophyll < 3 )


```

# Chloro plot
```{r}

chloro_plot <- ggplot() +
  geom_raster(data = final_chloro, aes(y=latitude, x=longitude, fill = chlorophyll), 
              show.legend = T, 
              interpolate = T, 
              hjust = 1, 
              vjust = 1) +
  scale_fill_viridis_c(option = "turbo") +
  theme_light() +
  labs(x = "Longitude",
       y = "",
       fill = "Chlorophyll
(mg m^-3)") +
  theme(panel.grid = element_line(F),
        panel.background = element_rect(fill = "grey70"),
        legend.position = "bottom",
        axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        plot.margin = unit(c(0,0.8,0,0), "inches")) +
  scale_x_continuous(breaks = c(-120.5, -120.25, -120, -119.75, -119.5, -119.25, -119), 
                     expand = c(0,0)) +
  scale_y_continuous(breaks = c(33.25, 33.5, 33.75, 34, 34.25, 34.5), 
                                expand = c(0,0), 
                     position = "right") +
  geom_sf(data = cha, color = "black", fill = "grey60") +
  geom_sf(data = ca, color = "black", fill = "grey60") +
  coord_sf(xlim = c(-120.69999999999999, -118.75), ylim = c(33.1125, 34.9), expand = FALSE)
  
  
```  
# Final Product
```{r}
grid.arrange(sst_plot, chloro_plot, ncol=2)
```




