---
title: "Houston Ice Storm"
description: |
  In my Spatial Analysis Class we calculated a rough estimate of the number of people who lost power from the 2021 Houston Ice Storm.
author:
  - name: Jake Eisaguirre
    url: {}
date: 2021-11-08
output:
  distill::distill_article:
    self_contained: false
---


# Impact of the February Storm on Power in Huston, TX

The extreme weather event seen in Houston, TX in February 2021 caused major blackouts due to a power grid failure. Using satellite images, the extent of the blackout can be measured by comparing the night lights before and after the storm.

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      include = TRUE)
```

```{r}
library(sf)
library(tidyverse)
library(stars)
library(rgdal)
library(stringr)
library(raster)
library(dplyr)
library(patchwork)
library(ggsn)
library(rosm)
library(tmap)
library(here)
library(rnaturalearth)

```

### Function to load the DNB dataset from VNP46A1 granules

```{r}
read_dnb <- function(file_name) {
  dataset_name <- "//HDFEOS/GRIDS/VNP_Grid_DNB/Data_Fields/DNB_At_Sensor_Radiance_500m"

  h_string <- gdal_metadata(file_name)[199]
  v_string <- gdal_metadata(file_name)[219]
  
  tile_h <- as.integer(str_split(h_string, "=", simplify = TRUE)[[2]])
  tile_v <- as.integer(str_split(v_string, "=", simplify = TRUE)[[2]])

  west <- (10 * tile_h) - 180
  north <- 90 - (10 * tile_v)
  east <- west + 10
  south <- north - 10

  delta <- 10 / 2400

  dnb <- read_stars(file_name, sub = dataset_name)

  st_crs(dnb) <- st_crs(4326)
  st_dimensions(dnb)$x$delta <- delta
  st_dimensions(dnb)$x$offset <- west
  st_dimensions(dnb)$y$delta <- -delta
  st_dimensions(dnb)$y$offset <- north
  
  return(dnb)
}
```

### Add data to dnb function

```{r}
feb07_5<- "data/VNP46A1.A2021038.h08v05.001.2021039064328.h5"
dnb07_5<- read_dnb(file_name = feb07_5)

feb07_6<- "data/VNP46A1.A2021038.h08v06.001.2021039064329.h5"
dnb07_6<- read_dnb(file_name = feb07_6)

feb16_5<- "data/VNP46A1.A2021047.h08v05.001.2021048091106.h5"
dnb16_5<- read_dnb(file_name = feb16_5)

feb16_6<- "data/VNP46A1.A2021047.h08v06.001.2021048091105.h5"
dnb16_6<- read_dnb(file_name = feb16_6)
```

### Combine based on date

```{r}
#feb 07 data
combined07 <- st_mosaic(dnb07_5,dnb07_6)

plot(combined07)

```

```{r}
#feb 16 data
combined16 <- st_mosaic(dnb16_5, dnb16_6)

plot(combined16)

rm(dnb07_5,
   dnb07_6,
   dnb16_5,
   dnb16_6) #remove unneeded files


```

### Create a blackout mask

```{r}
difference <- (combined07 - combined16) > 200

difference[difference == FALSE] <- NA
```
The data from after the storm was subtracted from that of before the storm to show the difference in night light caused by the power outage. 

### Vectorize the mask

```{r}
blackout <- st_as_sf(difference)
```

### Fix broken polygons

```{r}
blackout_mask <- st_make_valid(blackout)

```

### Crop the map to ROI

```{r}


poly <- st_polygon(list(rbind(c(-96.5, 29), 
                              c(-96.5, 30.5), 
                              c(-94.5, 30.5), 
                              c(-94.5, 29), 
                              c(-96.5, 29))))

ROI <-st_sfc(poly, crs = 4326) 


cropped <- blackout_mask[ROI, op = st_intersects]



map <- st_transform(cropped, crs = st_crs(3083))
```

### Account for highways

```{r}
query <- "SELECT * FROM gis_osm_roads_free_1 
WHERE fclass='motorway'"
roads <- ("data/gis_osm_roads_free_1.gpkg")
highways <- st_read(roads, query = query)
highways <- st_transform(highways, crs = st_crs(3083))

hwy_buffer200m <- st_buffer(highways, dist = 200)
hwy_dissolve <- st_union(hwy_buffer200m)
```

### Buildings Data

```{r}
query_1 <- "SELECT * FROM gis_osm_buildings_a_free_1 
WHERE (type IS NULL AND name IS NULL) OR 
type in ('residential', 'apartments', 'house', 'static_caravan', 'detached')"
buildings <- "data/gis_osm_buildings_a_free_1.gpkg"
houses <- st_read(buildings, query = query_1)
houses <- st_transform(houses, crs = st_crs(3083))
```

### Cencus Tract Data

```{r}
cencus <- "data/ACS_2019_5YR_TRACT_48_TEXAS.gdb"

acs_geoms <- st_read(cencus,layer = "ACS_2019_5YR_TRACT_48_TEXAS")

acs_income <- st_read(cencus,layer = "X19_INCOME")

median_income <- acs_income[c("GEOID", "B19013e1")]

median_income <- median_income %>% 
  rename(median_income = B19013e1)


```

### Merge data sets

```{r}
remove_highways <- st_difference(map, hwy_dissolve)

tot_num_of_houses_no_power <- houses[remove_highways, op = st_intersects]  

nrow(tot_num_of_houses_no_power)

texas_map <- ne_countries(country = 'united states of america', 
                         scale = 'large', returnclass = 'sf')

query_1 = "
  SELECT fclass, geom
  FROM gis_osm_roads_free_1
  WHERE fclass in ('motorway', 'motorway_link', 'primary', 'primary_link')"
highways2 = st_read("data/gis_osm_roads_free_1.gpkg", query = query)


Houston_map <- ggplot() +
  geom_sf(data = texas_map, fill = NA, color = 'black') +
  geom_sf(data = houses,color = "grey") +
  geom_sf(data = tot_num_of_houses_no_power, color = 'black')+
  geom_sf(data = highways2, color = "grey20") +
  theme_minimal() +
  ylab("Latitude") +
  xlab("Longitude") +
  ggtitle("Impact of 2021 Storm on Residential Houses in Houston") +
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1)) + 
  coord_sf(xlim = c(-94.5, -96.5), ylim = c(29, 30.5), expand = F) +
  north(x.min = -94.5, x.max = -96.5, y.min = 29, y.max = 30.5, scale = 0.15, location = "topright") +
  scalebar(x.min = -94.5, x.max = -96.5, y.min = 29, y.max = 30.5, dist = 25, dist_unit = "km", 
           transform = T, 
           st.bottom = F,
           location = "bottomleft",
           anchor = c(x = -95.2, y = 29.01))
Houston_map


```
Grey layer is the total number of residential houses in Houston and the black layer is the total number of residential houses that lost power from the 2021 ice storm. Dark gray lines are major road ways and dark black lines are Texas coastline. 

Due to the light pollution caused by highways, highways were removed from the dataset to not be included in the change of light. Comparing the night lights of February 16, 2021 to February 7, 2021, it appears that 157,411 houses were impacted by the power grid failure.

### Socioeconomic Impact

```{r}

socio_eco <- left_join(acs_geoms, median_income, by = c("GEOID_Data" = "GEOID"))

Residential_income <- ggplot(data = socio_eco) +
  geom_histogram(aes(x = median_income)) +
  theme_classic() +
  geom_vline(xintercept = 55771,
             col = 'red',
             lwd = 1)+
  annotate("text",
           x = 150000,
           y = 350,
           label = paste("Median = $55,771.00"),
           col = 'red')+
  geom_vline(xintercept = 64120.33,
             col = 'blue',
             lwd = 1)+
  annotate("text",
           x = 150000,
           y = 400,
           label = paste("Mean = $64,120.33"),
           col = 'blue') +
  ggtitle("Median Income of Residential Buildings in Houston")#all of houston

Residential_income



socio_eco <- st_transform(socio_eco, crs = st_crs(3083))
lost_power_income <- socio_eco[tot_num_of_houses_no_power, op = st_intersects]


Blackout_income <- ggplot(data = lost_power_income) +
  geom_histogram(aes(x = median_income)) +
  theme_classic() +
  geom_vline(xintercept = 60414.5,
             col = 'red',
             lwd = 1)+
  annotate("text",
           x = 150000,
           y = 50,
           label = paste("Median = $60,414.50"),
           col = 'red')+
  geom_vline(xintercept = 71244.88,
             col = 'blue',
             lwd = 1)+
  annotate("text",
           x = 150000,
           y = 75,
           label = paste("Mean = $71,244.88"),
           col = 'blue') +
  ggtitle("Median Income of Residential Buildings that Lost Power in Houston")#ice storm

Blackout_income
 

na_median <- na.omit(median_income)
median(na_median$median_income)
mean(na_median$median_income)

na_median_power <- na.omit(lost_power_income)
median(na_median_power$median_income)
mean(na_median_power$median_income)


```
Median income of all of Houston = $55,771.00
Mean income of all of Houston = $64,120.33
Median income of those that lost power in Houston = $60,414.50
Mean income of those that lost power in Houston = $71,244.88

Median income is not a factor in determining if a house will be affected by loss of power. There was a slightly higher median income with the houses that lost power. 

